---
title: "facility-characteristics"
author: "UWCHR"
date: "2/12/2021"
output:
  html_document: default
  pdf_document: default
---

Analyzing detention facility population by facility characteristic: "AS NEEDED" facilities; DMCP-authorized facilities; "Over-72" hour facilities. 

Starting with annual ADP. Recall that ADP does not equate to number of people passing through facility. Check book-ins along same lines.

To-do:
- [ ] Population measures by contract type/jurisdiction (may require light cleaning)
- [ ] Book-ins, initial book-ins measures by facility characteristic
- [ ] Compare measures for three main categories of facilities: Over-72 hour DMCP; Under-72 hour + non-DMCP; ORR
- [ ] Compare measures for sampled facilities versus non-sampled facilities

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# A few of these redundant
library(tidyverse)
library(here)
library(lubridate)
library(skimr)
library(tigris)
library(sf)
library(yaml)
library(here)
library(assertr)
library(ggrepel)

```

```{r load_files, echo=FALSE, message=FALSE, warning=FALSE,}

facilities <- read_delim(
  here::here('analyze', 'input', 'facil-list.csv.gz'), 
  delim = "|") 

dmcp <- read_delim(
  here::here('analyze', 'input', 'dmcp-detailed.csv.gz'), 
  delim = "|") 

```

```{r facil_data_setup_and_filter, echo=FALSE, message=FALSE, warning=FALSE,}

# Ideally specify field type on import
facilities <- facilities %>%
  mutate(detloc = as.factor(detloc))

facilities <- facilities %>%
  mutate(state = as.factor(state))

facilities <- facilities %>%
  mutate(city = as.factor(city))

facilities <- facilities %>%
  mutate(zip = as.factor(zip))

facilities <- facilities %>%
  mutate(aor = as.factor(aor))

facilities <- facilities %>%
  mutate(type = as.factor(type))

dmcp <- dmcp %>%
  mutate(state = as.factor(state),
         aor = as.factor(aor),
         detloc = as.factor(detloc))

dmcp$over_under_72 <- as.factor(dmcp$over_under_72)

# We don't actually want to filter by facility type here except ORR

facilities <- facilities %>%
  filter(type != "ORR")

# dmcp <- dmcp %>% 
#   filter(over_under_72 == "Over 72")

# dmcp_as_needed <- dmcp %>% 
#   filter(capacity == "AS NEEDED")

# facilities <- facilities %>% 
#   filter(detloc %in% dmcp$detloc)


# Concatenate facility city and state columns
facilities <- facilities %>% 
  unite("city", c(city,state), sep = ", ", remove = FALSE, na.rm = FALSE) %>%
  mutate(city = as.factor(city))

```

"AS NEEDED" facilities, though more common overall than facilities with known ICE capacity, held a smaller and diminishing portion of detained people by total average ADP over the period from FY2009 - FY2018; most book-ins also at facilities with known ICE capacity (note total book-ins may double-count people due to facilty transfers):

```{r as_needed_adp_natl, echo=FALSE, message=FALSE, warning=FALSE,}

facilities$as_needed <- facilities$capacity == 'AS NEEDED'

as_needed_adp <- facilities %>% 
  filter(!is.null(capacity)) %>% 
  group_by(as_needed) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

as_needed_bookins <- facilities %>% 
  filter(!is.null(capacity)) %>% 
  group_by(as_needed) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

as_needed_adp$fy <- as_needed_adp$fy %>% str_replace_all('[aA-zZ]', '')
as_needed_adp$fy <- paste('20', as_needed_adp$fy, sep='')
as_needed_adp$fy <- as.integer(as_needed_adp$fy) 

as_needed_bookins$fy <- as_needed_bookins$fy %>% str_replace_all('[aA-zZ]', '')
as_needed_bookins$fy <- paste('20', as_needed_bookins$fy, sep='')
as_needed_bookins$fy <- as.integer(as_needed_bookins$fy) 

# data <- left_join(as_needed_adp, as_needed_bookins, by=c('as_needed', 'fy'))

p1 <- ggplot(data = as_needed_adp,
       aes(x = fy, 
           y = adp, 
           group = as_needed, 
           color = as_needed)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total ADP") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly ADP by 'AS NEEDED' capacity status")

p1

p2 <- ggplot(data = as_needed_bookins,
       aes(x = fy, 
           y = bookins, 
           group = as_needed, 
           color = as_needed)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total bookins") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly total bookins by 'AS NEEDED' capacity status")

p2

```

'AS NEEDED' yearly ADP per AOR shows varied distributions and trends; 'AS NEEDED' ADP tends to be mostly stable year to year, while facilities with known capacity show more notable trends. Note divergence in New Orleans AOR 'NOL'. Total book-ins distributions vary, from AORs with majority booked in to facilities with known capacity, to AORs with roughly equal trends (suggestive of intra-AOR transfers?), to AORs with more book-ins at 'AS NEEDED' facilities:

```{r as_needed_adp_aor, echo=FALSE, message=FALSE, warning=FALSE,}

as_needed_adp_aor <- facilities %>% 
  filter(!is.null(capacity)) %>% 
  group_by(aor, as_needed) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

as_needed_bookins_aor <- facilities %>% 
  filter(!is.null(capacity)) %>% 
  group_by(aor, as_needed) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

as_needed_adp_aor$fy <- as_needed_adp_aor$fy %>% str_replace_all('[aA-zZ]', '')
as_needed_adp_aor$fy <- paste('20', as_needed_adp_aor$fy, sep='')
as_needed_adp_aor$fy <- as.integer(as_needed_adp_aor$fy) 

as_needed_bookins_aor$fy <- as_needed_bookins_aor$fy %>% str_replace_all('[aA-zZ]', '')
as_needed_bookins_aor$fy <- paste('20', as_needed_bookins_aor$fy, sep='')
as_needed_bookins_aor$fy <- as.integer(as_needed_bookins_aor$fy) 

p1 <- ggplot(data = as_needed_adp_aor,
       aes(x = fy, 
           y = adp, 
           group = as_needed, 
           color = as_needed)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total ADP") +
  theme_bw() + 
  facet_wrap(~ aor) + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly total ADP by 'AS NEEDED' capacity status")

p1

p2 <- ggplot(data = as_needed_bookins_aor,
       aes(x = fy, 
           y = bookins, 
           group = as_needed, 
           color = as_needed)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total bookins") +
  theme_bw() + 
  facet_wrap(~ aor) + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly total bookins by 'AS NEEDED' capacity status")

p2

```

Most AORs held low proportion of population in 'AS NEEDED' facilities. Wide spread in proportion of bookins in 'AS NEEDED' facilities.

```{r as_needed_adp_aor_proportion, echo=FALSE, message=FALSE, warning=FALSE,}

as_needed_adp_aor <- facilities %>% 
  filter(!is.null(capacity)) %>% 
  group_by(aor, as_needed) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>% 
    pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  ) %>% 
  pivot_wider(
    names_from = 'as_needed',
    values_from = 'adp'
  )

as_needed_bookins_aor <- facilities %>% 
  filter(!is.null(capacity)) %>% 
  group_by(aor, as_needed) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  ) %>% 
  pivot_wider(
    names_from = 'as_needed',
    values_from = 'bookins'
  )

as_needed_adp_aor$fy <- as_needed_adp_aor$fy %>% str_replace_all('[aA-zZ]', '')
as_needed_adp_aor$fy <- paste('20', as_needed_adp_aor$fy, sep='')
as_needed_adp_aor$fy <- as.integer(as_needed_adp_aor$fy) 

as_needed_bookins_aor$fy <- as_needed_bookins_aor$fy %>% str_replace_all('[aA-zZ]', '')
as_needed_bookins_aor$fy <- paste('20', as_needed_bookins_aor$fy, sep='')
as_needed_bookins_aor$fy <- as.integer(as_needed_bookins_aor$fy) 

as_needed_adp_aor$proportion_as_needed <- as_needed_adp_aor$'TRUE' / as_needed_adp_aor$'FALSE'

as_needed_bookins_aor$proportion_as_needed <- as_needed_bookins_aor$'TRUE' / as_needed_bookins_aor$'FALSE'

p1 <- as_needed_adp_aor %>% 
  filter(aor != 'BAL') %>%
  ggplot(
       aes(x = fy, 
           y = proportion_as_needed, 
           group = aor, 
           color = aor)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Proportion ADP") +
  theme_bw() +  
  facet_wrap(~ aor) +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Proportion ADP 'AS NEEDED'", caption="*FY2019 data through Aug. 2019, 'BAL' excluded")

p1

p2 <- as_needed_bookins_aor %>% 
  filter(aor != 'BAL') %>%
  ggplot(
       aes(x = fy, 
           y = proportion_as_needed, 
           group = aor, 
           color = aor)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Proportion bookins") +
  theme_bw() +  
  facet_wrap(~ aor) +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Proportion bookins 'AS NEEDED'", caption="*FY2019 data through Aug. 2019, 'BAL' excluded")

p2

```

DMCP-authorized facilities held/booked more and increasing number of detained people:

```{r dmcp_adp_natl, echo=FALSE, message=FALSE, warning=FALSE,}

facilities$dmcp <- facilities$detloc %in% dmcp$detloc

dmcp_adp <- facilities %>% 
  group_by(dmcp) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

dmcp_bookins <- facilities %>% 
  group_by(dmcp) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

dmcp_adp$fy <- dmcp_adp$fy %>% str_replace_all('[aA-zZ]', '')
dmcp_adp$fy <- paste('20', dmcp_adp$fy, sep='')
dmcp_adp$fy <- as.integer(dmcp_adp$fy) 

dmcp_bookins$fy <- dmcp_bookins$fy %>% str_replace_all('[aA-zZ]', '')
dmcp_bookins$fy <- paste('20', dmcp_bookins$fy, sep='')
dmcp_bookins$fy <- as.integer(dmcp_bookins$fy) 

p1 <- ggplot(data = dmcp_adp,
       aes(x = fy, 
           y = adp, 
           group = dmcp, 
           color = dmcp)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total ADP") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly ADP by DMCP status")

p1

p2 <- ggplot(data = dmcp_bookins,
       aes(x = fy, 
           y = bookins, 
           group = dmcp, 
           color = dmcp)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total bookins") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly bookins by DMCP status")

p2

```

Varied distributions and trends of detained population (ADP) held in DMCP-authorized facilities; most AOR similar to national trend of greater & increasing population in DMCP facilities.

```{r dmcp_adp_aor, echo=FALSE, message=FALSE, warning=FALSE,}

dmcp_adp_aor <- facilities %>% 
  group_by(aor, dmcp) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

dmcp_bookins_aor <- facilities %>% 
  group_by(aor, dmcp) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

dmcp_adp_aor$fy <- dmcp_adp_aor$fy %>% str_replace_all('[aA-zZ]', '')
dmcp_adp_aor$fy <- paste('20', dmcp_adp_aor$fy, sep='')
dmcp_adp_aor$fy <- as.integer(dmcp_adp_aor$fy) 

dmcp_bookins$fy <- dmcp_bookins$fy %>% str_replace_all('[aA-zZ]', '')
dmcp_bookins$fy <- paste('20', dmcp_bookins$fy, sep='')
dmcp_bookins$fy <- as.integer(dmcp_bookins$fy) 

p1 <- ggplot(data = dmcp_adp_aor,
       aes(x = fy, 
           y = adp, 
           group = dmcp, 
           color = dmcp)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total ADP") +
  theme_bw() + 
  facet_wrap(~ aor) + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly total ADP by DMCP capacity status")

p1

p2 <- ggplot(data = dmcp_bookins_aor,
       aes(x = fy, 
           y = bookins, 
           group = dmcp, 
           color = dmcp)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total bookins") +
  theme_bw() + 
  facet_wrap(~ aor) + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly total bookins by DMCP capacity status")

p2

```

Greater and slightly-increasing proportion of ADP in "Over 72" facilities. More bookins in "Over 72" facilities as well.

Should do some more checking for what facilities with null values in `over_under_72` represent:

```{r over_72_adp_natl, echo=FALSE, message=FALSE, warning=FALSE,}

facilities$over_72 <- facilities$over_under_72 == "Over 72"

over_72_adp <- facilities %>% 
  filter(!is.null(over_under_72)) %>%
  group_by(over_72) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

over_72_bookins <- facilities %>% 
  filter(!is.null(over_under_72)) %>%
  group_by(over_72) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

over_72_adp$fy <- over_72_adp$fy %>% str_replace_all('[aA-zZ]', '')
over_72_adp$fy <- paste('20', over_72_adp$fy, sep='')
over_72_adp$fy <- as.integer(over_72_adp$fy) 

over_72_bookins$fy <- over_72_bookins$fy %>% str_replace_all('[aA-zZ]', '')
over_72_bookins$fy <- paste('20', over_72_bookins$fy, sep='')
over_72_bookins$fy <- as.integer(over_72_bookins$fy) 

p1 <- ggplot(data = over_72_adp,
       aes(x = fy, 
           y = adp, 
           group = over_72, 
           color = over_72)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total ADP") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly ADP by Over-72 hour status")

p1

p2 <- ggplot(data = over_72_bookins,
       aes(x = fy, 
           y = bookins, 
           group = over_72, 
           color = over_72)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total bookins") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly bookins by Over-72 hour status")

p2

```

Note expansion of "Over 72" ADP in Atlanta 'ATL', San Antonio 'SNA' AORs; more booked in at "Under 72" facilities in DAL, PHO, SND: 

```{r over_72_adp_aor, echo=FALSE, message=FALSE, warning=FALSE,}

over_72_adp_aor <- facilities %>% 
  filter(!is.null(over_under_72)) %>% 
  group_by(aor, over_72) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

over_72_bookins_aor <- facilities %>% 
  filter(!is.null(over_under_72)) %>% 
  group_by(aor, over_72) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

over_72_adp_aor$fy <- over_72_adp_aor$fy %>% str_replace_all('[aA-zZ]', '')
over_72_adp_aor$fy <- paste('20', over_72_adp_aor$fy, sep='')
over_72_adp_aor$fy <- as.integer(over_72_adp_aor$fy) 

over_72_bookins_aor$fy <- over_72_bookins_aor$fy %>% str_replace_all('[aA-zZ]', '')
over_72_bookins_aor$fy <- paste('20', over_72_bookins_aor$fy, sep='')
over_72_bookins_aor$fy <- as.integer(over_72_bookins_aor$fy) 

p1 <- ggplot(data = over_72_adp_aor,
       aes(x = fy, 
           y = adp, 
           group = over_72, 
           color = over_72)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total ADP") +
  theme_bw() + 
  facet_wrap(~ aor) + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly total ADP by Over-72 hour status")

p1

p2 <- ggplot(data = over_72_bookins_aor,
       aes(x = fy, 
           y = bookins, 
           group = over_72, 
           color = over_72)) + 
  geom_line() +
  xlab("Fiscal Year") +
  ylab("Total bookins") +
  theme_bw() + 
  facet_wrap(~ aor) + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5)) +
  labs(title="Yearly total bookins by Over-72 hour status")

p2

```

```{r overall, echo=FALSE, message=FALSE, warning=FALSE,}

dmcp <- dmcp %>%
  filter(over_under_72 == "Over 72")

dmcp_as_needed <- dmcp %>%
  filter(capacity == "AS NEEDED")

facilities$detloc <- as.character(facilities$detloc)

facilities$sample <- facilities %>%
  filter(detloc %in% dmcp$detloc)

```
