---
title: "facility-type"
author: "UWCHR"
date: "2/19/2021"
output: html_document
---

Descriptive analysis of ICE detention statistics based on facility type.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# A few of these redundant
library(tidyverse)
library(here)
library(lubridate)
library(skimr)
library(tigris)
library(sf)
library(yaml)
library(here)
library(assertr)
library(ggrepel)
library(ggalt)

```

```{r load_files, echo=FALSE, message=FALSE, warning=FALSE,}

facilities <- read_delim(
  here::here('analyze', 'input', 'facil-list.csv.gz'), 
  delim = "|") 

dmcp <- read_delim(
  here::here('analyze', 'input', 'dmcp-detailed.csv.gz'), 
  delim = "|") 

```

```{r facil_data_setup_and_filter, echo=FALSE, message=FALSE, warning=FALSE}

# Ideally specify field type on import
facilities <- facilities %>%
  mutate(detloc = as.factor(detloc))

facilities <- facilities %>%
  mutate(state = as.factor(state))

facilities <- facilities %>%
  mutate(city = as.factor(city))

facilities <- facilities %>%
  mutate(zip = as.factor(zip))

facilities <- facilities %>%
  mutate(aor = as.factor(aor))

facilities <- facilities %>%
  mutate(type = as.factor(type))

dmcp <- dmcp %>%
  mutate(state = as.factor(state),
         aor = as.factor(aor),
         detloc = as.factor(detloc))

dmcp$over_under_72 <- as.factor(dmcp$over_under_72)

facilities <- facilities %>%
  filter(type != "ORR")

facilities$as_needed <- facilities$capacity == 'AS NEEDED'

facilities$dmcp <- facilities$detloc %in% dmcp$detloc

facilities$over_72 <- facilities$over_under_72 == "Over 72"

sample <- dmcp %>%
  filter(over_under_72 == "Over 72")

facilities$sample <-facilities$detloc %in% sample$detloc

# Concatenate facility city and state columns
facilities <- facilities %>% 
  unite("city", c(city,state), sep = ", ", remove = FALSE, na.rm = FALSE) %>%
  mutate(city = as.factor(city))

```

By ADP, most detained people held in IGSA, CDF, USMS IGA facilities:

```{r type_adp, echo=FALSE, message=FALSE, warning=FALSE}

data <- facilities %>% 
  group_by(type) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

data$fy <- data$fy %>% str_replace_all('[aA-zZ]', '')
data$fy <- paste('20', data$fy, sep='')
data$fy <- as.integer(data$fy) 


p1 <- data %>%
  ggplot(
       aes(x = fy, 
           y = adp, 
           group = type, 
           color = type)) + 
  geom_line() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p1

```

Most detained people booked in to "Other", IGSA, USMS IGA facilities. "Other" includes hold rooms, staging facilities (look more closely at this category):

```{r type_bookins, echo=FALSE, message=FALSE, warning=FALSE}

data <-  facilities %>% 
  group_by(type) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

data$fy <- data$fy %>% str_replace_all('[aA-zZ]', '')
data$fy <- paste('20', data$fy, sep='')
data$fy <- as.integer(data$fy) 

p1 <- data %>%
  ggplot(
       aes(x = fy, 
           y = adp, 
           group = type, 
           color = type)) + 
  geom_line() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p1

```

Similar trends for initial bookins:

```{r type_initial_bookins, echo=FALSE, message=FALSE, warning=FALSE}

data <-  facilities %>% 
  group_by(type) %>% 
  summarize_at(vars(contains("initial"), -contains("adp"), ), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

data$fy <- data$fy %>% str_replace_all('[aA-zZ]', '')
data$fy <- paste('20', data$fy, sep='')
data$fy <- as.integer(data$fy) 

p1 <- data %>%
  ggplot(
       aes(x = fy, 
           y = adp, 
           group = type, 
           color = type)) + 
  geom_line() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p1

```

Scatter plot of yearly observations of ADP, bookins per detention facility. Shows clear groupings of specific kinds of facilities, likely log linear relationship overall between ADP, bookings.

Interesting to consider which kinds of facilities play which role in different AORs, for example, see plot 3: high-ADP, high-bookins quadrant represented by contract detention facilities in SEA, DEN; service processing centers in BUF; BOP and IGSA facilities in NYC; IGSA facilities tend to play flexible role in many AORs.

```{r type_adp_bookins, echo=FALSE, message=FALSE, warning=FALSE}

data_adp <- facilities %>%
  select(c('detloc', 'aor', 'type', contains('adp'), -contains('bookin'))) %>% 
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

data_adp$fy <- data_adp$fy %>% str_replace_all('[aA-zZ]', '')
data_adp$fy <- paste('20', data_adp$fy, sep='')
data_adp$fy <- as.integer(data_adp$fy) 

data_bookins <- facilities %>%
  select(c('detloc', 'aor', 'type', contains("bookins"), -contains("adp"), -contains("initial"))) %>% 
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

data_bookins$fy <- data_bookins$fy %>% str_replace_all('[aA-zZ]', '')
data_bookins$fy <- paste('20', data_bookins$fy, sep='')
data_bookins$fy <- as.integer(data_bookins$fy) 

data <- merge(data_adp, data_bookins, by=c('detloc', 'aor', 'type', 'fy'))

p1 <- data %>% 
  ggplot(aes(
    x = log(adp),
    y = log(bookins),
    color = type,
    shape = factor(fy)
  )) +
  geom_point() +
  geom_encircle(aes(group=type)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p1

p2 <- data %>% 
  ggplot(aes(
    x = log(adp),
    y = log(bookins),
    color = type,
    shape = factor(fy)
  )) +
  geom_point() +
  facet_wrap(~aor) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p2


p3 <- data %>% 
  ggplot(aes(
    x = log(adp),
    y = log(bookins),
    color = type,
    shape = factor(fy)
  )) +
  geom_encircle(aes(group=type, fill=type), alpha=.5) +
  facet_wrap(~aor) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p3
```

ADP and bookins trends per AOR. Note AORs where most active facility type has changed: IGSAs become more prominent in SNA, SFR, WAS; contract detention facilities become more prominent in SND.

```{r per_aor, echo=FALSE, message=FALSE, warning=FALSE}

data_bookins <- facilities %>%
  group_by(aor, type) %>% 
  summarize_at(vars(contains("bookins"), -contains("adp"), -contains("initial")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "bookins",
    values_drop_na = TRUE
  )

data_adp <- facilities %>%
  group_by(aor, type) %>% 
  summarize_at(vars(ends_with("adp"), -contains("bookin")), sum) %>%
  pivot_longer(
    cols = starts_with("fy"),
    names_to = "fy",
    values_to = "adp",
    values_drop_na = TRUE
  )

data_bookins$fy <- data_bookins$fy %>% str_replace_all('[aA-zZ]', '')
data_bookins$fy <- paste('20', data_bookins$fy, sep='')
data_bookins$fy <- as.integer(data_bookins$fy) 

data_adp$fy <- data_adp$fy %>% str_replace_all('[aA-zZ]', '')
data_adp$fy <- paste('20', data_adp$fy, sep='')
data_adp$fy <- as.integer(data_adp$fy) 

p1 <- data_adp %>%
  ggplot(
       aes(x = fy, 
           y = adp, 
           group = type, 
           color = type)) + 
  geom_line() +
  facet_wrap(~aor) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p1

p2 <- data_bookins %>%
  ggplot(
       aes(x = fy, 
           y = bookins, 
           group = type, 
           color = type)) + 
  geom_line() +
  facet_wrap(~aor) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = .5))

p2

```